---
title: "ass3"
author: "Tiffany Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions: 16/16

## Q 1 5.54 pg 215

a. 0.2827 or 0.2888
```{r}
fish <- read.csv("PHISHING.csv")

mean(fish$INTTIME)
# P(Y>120)
mean(fish$INTTIME > 120)
# a. 28.84%
```

$$
P(Y>120) = \int_{120}^{\infty} \frac{e^{-y/95}}{95} dy = -e^{-y/95}|_{120}^{\infty} = -e^{\infty/95} + e^{-120/95} = 0 + 0.282759689796 = 0.282759689796

$$

b. yes, the data follows the exponential distribution with a mean of 95.5237 and sd of 91.5391. Both the mean and standard deviation are close to 95 meaning the data does appear to fllow an exponential distribution with $\beta = 95$

```{r}

# mean and sd
mean(fish$INTTIME)
sd(fish$INTTIME)

# summary for funsies
summary(fish$INTTIME)
```

## Q 2 5.56 pg 215

a. mean = 0.21, variance = 0.0147

$$
\alpha = 3 \newline
\beta = 0.07 \newline
\newline
\mu = \alpha\beta = (3)(0.07) = 0.21 \newline
\sigma^{2} = \alpha\beta^{2} = (3)(0.07)^{2} = 0.0147 \newline
$$

b. No, 0.60 falls outside of the 3 sd range when $\alpha=3$ and $\beta=0.07$. The previous gamma distribution no longer models the new maximum flood level distribution for the 4-year period.

$$
\sigma = \sqrt{\sigma^{2}} = \sqrt{0.0147} = 0.1212435565 \newline
\mu \pm 3\sigma = 0.21 \pm 3(0.1212435565) = 0.21 \pm 0.3637306695 = (-0.1537306695, 0.5737306695) \newline
$$

## Q 3 5.60 pg 216

a. meanA = 4, meanB = 4

$$
formula A: \alpha=2, \beta=2 \newline
formula B: \alpha=1, \beta=4 \newline
\mu_A = \alpha\beta=(2)(2)=4 \newline
\mu_B = \alpha\beta=(1)(4)=4 \newline
$$

b. variance A = 8, variance B = 16

$$
\sigma^{2}_{A} = \alpha\beta^{2} = (2)(2)^{2} = 8 \newline
\sigma^{2}_{B} = \alpha\beta^{2} = (1)(4)^{2} = 16 \newline
$$

c.

For A:

$$
P(Y<1)=\int_0^1 \frac{y^{2-1}e^{-y/2}}{2^2\Gamma(2)}dy = \int_0^1 \frac{ye^{-y/2}}{4}dy = \frac{1}{4} \int_0^1 2e^{-y/2}dy + \frac{1}{4} (-2e^{-y/2})|_0^1 \newline
= \frac{1}{4} \left[-2e^{-1/2} + 2(0)e^{0}\right] = \frac{1}{4} \left[-2e^{-1/2}(3) + 4\right] = 0.09020401045
$$

For B:

$$
P(Y<1)=\int_0^1 \frac{y^{1-1}e^{-y/4}}{4^1\Gamma(1)}dy = \int_0^1 \frac{ye^{-y/4}}{4}dy = \frac{1}{4} \int_0^1 4e^{-y/4}dy + \frac{1}{4} (-4e^{-y/4})|_0^1 \newline
= \frac{1}{4} \left[-4e^{-1/4} + 4(0)e^{0}\right] = \frac{1}{4} \left[-4e^{-1/4}(1) + 4\right] = 0.2211992169286
$$

Tear gas Formula B has a higher probability (0.2211 > 0.0902) of generating a human reaction in less than 1 minute. The probability of Formula B is higher than the probability of Formula A.

## Q 4 5.74 pg 219

a. about 0.6321

Using CDF for Weibull distribution :D
$$
F(y_0) = \int_{0}^{y_0}f(y)dy = \int_{0}^{y_0}\frac{\alpha}{\beta}y^{\alpha-1}e^{-y^{\alpha}/ \beta}dy = 1-e^{-z/\beta} = 1-e^{-y_{0}^{a}/ \beta} \newline
P(Y<2) = 1-e^{-2^{2}/4} = 1-e^{-1} = 0.6321205588
$$

b. mean = 1.7724, sd = 0.9264

mean & standard deviation
$$
\mu = \beta^{1/\alpha}\Gamma(\frac{\alpha+1}{\alpha}) = (4)^{1/2}(\Gamma(\frac{2+1}{2})) = 2\Gamma(1.5) = 1.7724538509 \newline
\sigma^2=\beta^{2/\alpha}\left[\Gamma(\frac{\alpha+2}{\alpha})-\Gamma^{2}(\frac{\alpha+1}{\alpha})\right] = (4)^{2/2}\left[\Gamma(2)-\Gamma^{2}(1.5) \right] = (4)(1-(.88623)^{2}) = 0.9264909866803887
$$

c.
$$
P(\mu-2\sigma \le Y \le \mu+2\sigma) = P(1.7724-2(0.9264) \le Y \le 1.7724+2(0.9264)) = P(-0.0804 \le Y \le 3.6712) \newline
$$

d. No, the probability is 0.1053, or about 11%

$$
F(y_0) = \int_{0}^{y_0}f(y)dy = \int_{0}^{y_0}\frac{\alpha}{\beta}y^{\alpha-1}e^{-y^{\alpha}/\beta}dy = 1-e^{-(y/\beta)^{\alpha}} \newline
P(Y>6) = 1 - F(6) = 1 - (1-e^{-(6/4)^{2}}) = e^{-9/4} = 0.10539922456186433

$$

## Q 5 5.84 pg 223

a. mean = 0.1818, variance = 0.0123

$$
\mu = \frac{\alpha}{\alpha+\beta} = \frac{2}{2+9} = \frac{2}{11} = 0.18181818 \newline
\sigma^{2}=\frac{\alpha\beta}{(\alpha+\beta)^{2}(\alpha+\beta+1)} = \frac{2(9)}{(2+9)^{2}(2+9+1)} = \frac{18}{1452} = 0.01239669421
$$

b. 0.04636 or 4%

$$
PDF: f(y)=\frac{y^{\alpha-1}(1-y)^{\beta-1}}{B(\alpha, \beta)} = \frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\Gamma(\beta)}y^{\alpha-1}(1-y)^{\beta-1}, 0 \le y \le 1 \newline
\alpha=2,\beta=9: f(y)=\frac{ \Gamma(2+9) }{ \Gamma(2)\Gamma(9) }y^{2-1}(1-y)^{9-1} = \frac{ \Gamma(11) }{ \Gamma(2)\Gamma(9) }y^{1}(1-y)^{8} = \frac{(10!)y(1-y)^{8}}{(1!)(8!)} = 90y(1-y)^{8} \newline
P(Y>0.40) = \int_{0.40}^{1}90y(1-y)^{8}dy = \int_{u=0.6}^{0}90(1-u)u^{8}(-du) = 90\int_{0}^{0.6}(u^{8}-u^{9})du=90\left[\frac{u^{9}}{9} - \frac{u^{10}}{10} \right]_{0}^{0.6} = 90(\frac{0.6^{9}}{9} - \frac{0.6^{10}}{10}) = 0.04636
$$

c. 0.2639 or 26%

$$
n=(\alpha+\beta-1)=2+9-1=10,p=0.10; P(Y \le 10) = F(0.10) \newline
F(p)=\int_{0}^{p}\frac{y^{\alpha-1}(1-y)^{\beta-1}}{B(\alpha, \beta)}dy = \sum_{y=\alpha}^{n}p(y) \newline
F(0.10)=\sum_{y=2}^{10}p(y) = 1-\sum_{y=0}^{1}p(y) = 1 - 0.7361 = 0.2639
$$


## Q 6 5.114 pg 232

a. alpha = 2, beta = 16 (2/16 is 1/8 yay)

b. mean = 3.5449, variance = 3.4335

$$
\mu=\beta^{1/\alpha}\Gamma(\frac{\alpha+1}{\alpha}) = 16^{1/2}\Gamma(\frac{3}{2}) = (4)\Gamma(1.5) = (4)(0.88623) = 3.54492 \newline
\sigma^{2}=\beta^{2/\alpha}\left[\Gamma(\frac{\alpha+2}{\alpha}) - \Gamma^{2}(\frac{\alpha+1}{\alpha}) \right] = 16^{2/2}\left[ \Gamma(\frac{2+2}{2})-\Gamma^{2}(\frac{2+1}{2}) \right] = (16)\left[\Gamma(2)-\Gamma^{2}(1.5)\right] = (16)(1-(0.88623)^{2}) = 3.433542194
$$

c. 0.1054 or 10%

$$
P(Y \ge 6) = 1 - P(Y<6) = 1-F(6) \newline
F(y_0) = \int_{0}^{y_0}f(y)dy = \int_{0}^{y_0}\frac{\alpha}{\beta}y^{\alpha-1}e^{-y^{\alpha}/\beta}dy \rightarrow F(6) = \int_{0}^{6}\frac{2}{16}y^{2-1}e^{-y^{2}/16}dy = \frac{1}{8}\int_{0}^{6}ye^{-y^{2}/16}dy = \frac{1}{8}\left[-8e^{-y^{2}/16}\right]_{0}^{6} = \frac{1}{8}\left[-8e^{-36/16}+8e^{0}\right] = \frac{1}{8}\left[8(1-e^{-9/4})\right] = 1-e^{-9/4} = 0.1054
$$


## Q 7 6.2 pg 239

a. We know 1 $\le$ X $\le$ 6 and 1 $\le$ Y $\le$ 6 also 36 outcomes is equally likely. Then the joint probability distribution is:

$$
p(x,y) = \frac{1}{36} \text{ for all } x,y \in \{1, 2, 3, 4, 5, 6\}
$$

b. Since both dice are independent, the marginal distribution for both are the same: $\frac{1}{6}$

$$
p_{1}(x)=\sum_{y=1}^{6}p(x,y) = \sum_{y=1}^{6}\frac{1}{36} = \frac{6}{36} = \frac{1}{6} \newline
p_{2}(x)=\sum_{y=1}^{6}p(x,y) = \sum_{y=1}^{6}\frac{1}{36} = \frac{6}{36} = \frac{1}{6} \newline
$$

c.Since both dice are independent, the conditional probability distribution for both are the same: $\frac{1}{6}$

$$
p_{1}(x,y)=\frac{p(x,y)}{p_2(y)} = \frac{\frac{1}{36}}{\frac{1}{6}} = \frac{1}{6} \newline
p_{2}(x,y)=\frac{p(x,y)}{p_1(x)} = \frac{\frac{1}{36}}{\frac{1}{6}} = \frac{1}{6} \newline
$$

d. Since the conditional probability for both $p_{1}(x,y)$ and $p_{2}(x,y)$ are equal to the joint probability distribution, $\frac{1}{36}$, the outcomes of the two dice are therefore statistically independent. 


## Q 8 6.4 pg 240

a.

$$
x=1,y=1 \rightarrow ID=2 \rightarrow p(1,1)=\frac{1}{7} \newline
x=1,y=2 \rightarrow ID=N/A \rightarrow p(1,2)=0 \newline
x=1,y=3 \rightarrow ID=N/A \rightarrow p(1,2)=0 \newline
x=2,y=1 \rightarrow ID=4,7 \rightarrow p(2,1)=\frac{2}{7} \newline
x=2,y=2 \rightarrow ID=N/A \rightarrow p(2,2)=0 \newline
x=2,y=3 \rightarrow ID=N/A \rightarrow p(2,3)=0 \newline
x=3,y=1 \rightarrow ID=1 \rightarrow p(3,1)=\frac{1}{7} \newline
x=3,y=2 \rightarrow ID=5,6 \rightarrow p(3,2)=\frac{2}{7} \newline
x=3,y=3 \rightarrow ID=3 \rightarrow p(3,3)=\frac{1}{7} \newline
$$

b.

$$
p_1(x)=\sum_{y}p(x,y) \newline
p_1(1)=\frac{1}{7}+0+0=\frac{1}{7} \newline
p_1(2)=0+0+\frac{2}{7}=\frac{2}{7} \newline
p_1(3)=\frac{1}{7}+\frac{2}{7}+\frac{1}{7}=\frac{4}{7} \newline
$$
c.

$$
p_2(y)=\sum_{x}p(x,y) \newline
p_2(1)=\frac{1}{7}+\frac{2}{7}+\frac{1}{7}=\frac{4}{7} \newline
p_2(2)=0+0+\frac{2}{7}=\frac{2}{7} \newline
p_2(3)=0+0+\frac{1}{7}=\frac{1}{7} \newline
$$

d. conditional distribution

$$
p_{2}(y|x)=\frac{p(x,y)}{p_{1}(x)} \newline
p_{2}(1|1)=\frac{p(1,1)}{p_{1}(1)}=\frac{\frac{1}{7}}{\frac{1}{7}}=1 \newline
p_{2}(2|1)=\frac{p(1,2)}{p_{1}(1)}=\frac{0}{\frac{1}{7}}=0 \newline
p_{2}(3|1)=\frac{p(1,3)}{p_{1}(1)}=\frac{0}{\frac{1}{7}}=0 \newline
p_{2}(1|2)=\frac{p(2,1)}{p_{1}(2)}=\frac{\frac{2}{7}}{\frac{2}{7}}=1 \newline
p_{2}(2|2)=\frac{p(2,2)}{p_{1}(2)}=\frac{0}{\frac{2}{7}}=0 \newline
p_{2}(3|2)=\frac{p(2,3)}{p_{1}(2)}=\frac{0}{\frac{2}{7}}=0 \newline
p_{2}(1|3)=\frac{p(3,1)}{p_{1}(3)}=\frac{\frac{1}{7}}{\frac{4}{7}}=\frac{1}{4} \newline
p_{2}(2|3)=\frac{p(3,2)}{p_{1}(3)}=\frac{\frac{2}{7}}{\frac{4}{7}}=\frac{2}{4} \newline
p_{2}(3|3)=\frac{p(3,3)}{p_{1}(3)}=\frac{\frac{1}{7}}{\frac{4}{7}}=\frac{1}{4} \newline
$$


## Q 9 6.12 pg 244

a.

$$
\text{probability density of X and Y: }f(x,y)=\frac{e^{-y/10}}{10y}, 0<y<x<2y \newline
f(y)=\int_{y}^{2y}\frac{e^{-y/10}}{10y}dx = \frac{e^{-y/10}}{10y}(2y-y)=\frac{e^{-y/10}}{10y}y=\frac{e^{-y/10}}{10} \text{ or } \frac{1}{10}e^{-y/10} \newline
$$
This is the exponential distribution, where $\lambda=\frac{1}{10}$.

b. The mean DOT estimate $E(Y)=\mu=\beta=10$ (thousands of dollhairs)

$$
E(Y)=\frac{1}{\lambda}=\frac{1}{\frac{1}{10}}=10 
$$

## Q 10 6.14 pg 245

a. solving for c

$$
\text{initial setup: } \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}f(x,y)dxdy = \int_{0}^{x}\int_{0}^{\infty}ce^{-x^{2}}dydx \newline
\text{inner integral: } \rightarrow \int_{0}^{x}ce^{-x^{2}}dy = ce^{-x{2}}(x) \newline
\text{plug into outer integral} \rightarrow \int_{0}^{\infty}cxe^{-x^{2}}dx = \int_{0}^{\infty}xe^{-x^{2}}dx=\frac{1}{2}\int_{0}^{\infty}e^{-u}du=\frac{1}{2} \newline
\text{solve for c: } \rightarrow \frac{c}{2} = 1 \rightarrow c=2 \newline
$$

b. marginal density

$$
\int_{-\infty}^{\infty}f_{1}(x)dx=2\int_{0}^{1}xdx=2\Big( \frac{x^{2}}{2} \Big)\Big]_{0}^{1}=1 \newline
$$

c. conditional density for Y given X is a uniform distribution over the interval 0 $\le$ Y $\le$ X

$$
f_{2}(y|x)=\frac{f(x,y)}{f_1(x)}=\frac{2e^{-x^{2}}}{2xe^{-x^{2}}}=\frac{1}{x}
$$

## Q 11 6.51 pg 253

$$
Cov(X,Y)=E(XY)-E(X)E(Y) = \sum_{x}\sum_{y}xyp(x,y) \newline
$$

Marginal distribution of X

$$
p_{1}(-1)=\frac{1}{12}+\frac{2}{12}+\frac{1}{12}=\frac{4}{12} \newline
p_{1}(0)=\frac{2}{12}+0+\frac{2}{12}=\frac{4}{12} \newline
p_{1}(1)=\frac{1}{12}+\frac{2}{12}+\frac{1}{12}=\frac{4}{12} \newline
E(X)=\sum_{x}xp(x) = (-1)\Big(\frac{4}{12}\Big)+(0)\Big(\frac{4}{12}\Big)+(1)\Big(\frac{4}{12}\Big) = -\frac{4}{12}+\frac{4}{12} = 0
$$

Marginal distribution of Y

$$
p_{1}(-1)=\frac{1}{12}+\frac{2}{12}+\frac{1}{12}=\frac{4}{12} \newline
p_{1}(0)=\frac{2}{12}+0+\frac{2}{12}=\frac{4}{12} \newline
p_{1}(1)=\frac{1}{12}+\frac{2}{12}+\frac{1}{12}=\frac{4}{12} \newline
E(Y)=\sum_{y}yp(y) = (-1)\Big(\frac{4}{12}\Big)+(0)\Big(\frac{4}{12}\Big)+(1)\Big(\frac{4}{12}\Big) = -\frac{4}{12}+\frac{4}{12} = 0
$$

Calculate E[XY]

$$
E[XY]=\sum_{x}\sum_{y}xyp(x,y) \newline
x=-1, y=-1 \rightarrow (-1)(-1)\Big(\frac{1}{12}\Big) = \frac{1}{12} \newline
x=-1, y=0 \rightarrow (-1)(0)\Big(\frac{2}{12}\Big) = 0 \newline
x=-1, y=1 \rightarrow (-1)(1)\Big(\frac{1}{12}\Big) = -\frac{1}{12} \newline
x=0, y=-1 \rightarrow (0)(-1)\Big(\frac{2}{12}\Big) = 0 \newline
x=0, y=0 \rightarrow (0)(0)\Big(0\Big) = 0 \newline
x=0, y=1 \rightarrow (0)(1)\Big(\frac{2}{12}\Big) = 0 \newline
x=1, y=-1 \rightarrow (1)(-1)\Big(\frac{1}{12}\Big) = -\frac{1}{12} \newline
x=1, y=0 \rightarrow (1)(0)\Big(\frac{2}{12}\Big) = 0 \newline
x=1, y=1 \rightarrow (1)(1)\Big(\frac{1}{12}\Big) = \frac{1}{12} \newline
E[XY] = \frac{1}{12} + 0 - \frac{1}{12} + 0 + 0 + 0 - \frac{1}{12} + 0 + \frac{1}{12} = 0
$$

Calculate Covariance:

$$
Cov(X,Y)=E(XY)-E(X)E(Y) = 0 - (0)(0) = 0
$$

Show that X and Y are dependent:

$$
p(x,y)\neq p_{1}(x)p_{2}(y) \newline
\text{check any pair (x=0, y=0 for example): } p(0,0)=0 \neq p_{1}(0)p_{2}(0) = \Big(\frac{4}{12}\Big)\Big(\frac{4}{12}\Big) = \frac{16}{144} \neq 0 \newline
\text{WRONGGGGG...yeah X and Y are dependent}
$$

We have shown that X and Y are dependent, and that Cov(X,Y)=0. :3


## Q 12 6.74 pg 269

a.

$$
E(\bar{Y})=\mu = \frac{1+3}{2} = 2
$$

b.

$$
Var(\bar{Y})=\frac{(3-1)^{2}}{12 * 60} = 0.005555555
$$

c. The central limit theorem states that under appropriate conditions the distribution of a normalized version of the sample mean will converge to a standard normal distribution. Given the large sample size of 60, the shape of the sampling distribution of $\bar{Y}$ will be approximately normal.

d.

$$
P(\bar{Y} < 2.5) - P(\bar{Y} < 1.5) = P\Big(Z < \frac{2.5-2}{\sqrt{0.005555555}}\Big) - P\Big(Z < \frac{1.5-2}{\sqrt{0.005555555}}\Big) = 1
$$
e.

$$
P(\bar{Y} > 2.2) = 1 - P(\bar{Y} < 2.2) = 1 - P\Big(Z < \frac{2.2-2}{\sqrt{0.005555555}}\Big) = 1 - P(Z < 2.6845) = 0.9963
$$

## Q 13 6.90 pg 273

a. 0.0015

$$
\mu=(20)(0.4)=8 \newline
\sigma=\sqrt{(20)(0.4)(0.6)}=\sqrt{4.8}= 2.19089 \newline
z=\frac{x-\mu}{\sigma}=\frac{(2-0.5)-8}{2.19} = -2.9680 \rightarrow P(Z<-2.97)=0.0015
$$

b.

$$
z=\frac{x-\mu}{\sigma}=\frac{(10+0.5)-8}{2.19} = 1.1415 \rightarrow P(Z<1.14) = 0.1271
$$

c.

$$
\text{a: }P(X<2) = P(X=0)+P(X=1) = 0.5-0.4985 = 0.0015 \newline
\text{b: }P(X>10) = P(X=11)+P(X=12)+P(X=13)+P(X=14)+P(X=15)+P(X=16)+P(X=17)+P(X=18)+P(X=19)+P(X=20) = 1-0.8725 = 0.1275
$$

## Q 14 7.108 pg 362

a. please i calculated everything else by hand,....i used r just once...please.....i beg you

```{r}
leadcop <- read.csv("LEADCOPP.csv")
x = leadcop$LEAD
alpha  =0.01
n = length(x)
mp = c(-1,1)
mean(x) + mp*qt(1-alpha/2,n-1 )*sd(x)/sqrt(n)
```

b. okay maybe i used r twice...but...i beg you...please...its3am....im tired
```{r}
x = leadcop$COPPER
alpha  =0.01
n = length(x)
mp = c(-1,1)
mean(x) + mp*qt(1-alpha/2,n-1 )*sd(x)/sqrt(n)
```

c. 
The confidence interval in part a signifies that we are 99% confident the mean lead level in drinking water is between 1.15 and 6.92. 
The confidence interval in part b signifies that we are 99% confident the mean copper level in drinking water is between -0.25 and 0.25.

d. Being 99% confident means that if we were to take many samples and construct confidence intervals for each sample, approximately 99% of those intervals would contain the true population mean.


## Q 15 7.114 pg 364

a.

```{r}
solrad <- read.csv("SOLARAD.csv")
summary(solrad$STJOS)
summary(solrad$IOWA)
```

$$
\alpha=0.05 \rightarrow \frac{\alpha}{2} = \frac{0.05}{2} = 0.025 \newline
\text{degrees of freedom: }n=1=7-1=6, t_{0.025}=2.447 \newline
\bar{d} \pm t_{\alpha/2} \frac{s}{\sqrt{n}} \rightarrow 198.0\pm2.447\frac{44.5}{\sqrt{7}} \rightarrow 198.0\pm41.16 \rightarrow (156.84,239.16)
$$

We are 95% confident that the mean difference in solar radiation between the two sites is between 156.84 and 239.16.


## Q 16 7.116 pg 364

```{r}
diazinon <- read.csv("DIAZINON.csv")
summary(diazinon$DAY)
summary(diazinon$NIGHT)

day <- c(5.4, 2.7, 34.2, 19.9, 2.4, 7.0, 6.1, 7.7, 18.4, 27.1, 16.9)
night <- c(24.3, 16.5, 47.2, 12.4, 24.0, 21.6, 104.3, 96.9, 105.3, 78.7, 44.6)

differences <- day - night
sum(differences) # use this number for sum of differences

differences2 <- (day-night)^2
sum(differences2) # use this number for sum of squared differences
```

a.

$$
\bar{d}=\frac{\sum d_{i}}{n} = \frac{-428}{11} = 38.909 \newline
s_{d}^{2} = \frac{\sum d_{i}^{2} - \frac{\Big( \sum d_{i}\Big)^{2} }{n} }{n-1} = \frac{30033.96 - \frac{(-428)^{2}}{11} }{11-1} = 1338.0869 \newline
s_{d} = \sqrt{s_{d}^{2}} = \sqrt{1338.0869} = 36.5799 \newline
\bar{d}\pm t_{\alpha/2}\frac{s_{d}}{\sqrt{n}} \rightarrow 38.909\pm1.812\frac{36.5799}{\sqrt{11}} \rightarrow -38.909\pm19.985 \rightarrow (-58.894, -18.924)
$$

We are 95% confident that the true mean difference in diazinon levels between day and night is between -58.894 and -18.924.

b. We assume the population of differences is normal

c. Since the confidence interval does not contain 0, there is evidence of a difference in the mean diazinon residue between day and night. Since the interval only contains negative numbers, the night mean is greater than the day mean.



