---
title: "ass3"
author: "Tiffany Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions: 16/16

## Q 1 5.54 pg 215

a. 0.2827 or 0.2888
```{r}
fish <- read.csv("PHISHING.csv")

mean(fish$INTTIME)
# P(Y>120)
mean(fish$INTTIME > 120)
# a. 28.84%
```

$$
P(Y>120) = \int_{120}^{\infty} \frac{e^{-y/95}}{95} dy = -e^{-y/95}|_{120}^{\infty} = -e^{\infty/95} + e^{-120/95} = 0 + 0.282759689796 = 0.282759689796

$$

b. yes, the data follows the exponential distribution with a mean of 95.5237 and sd of 91.5391. Both the mean and standard deviation are close to 95 meaning the data does appear to fllow an exponential distribution with $\beta = 95$

```{r}

# mean and sd
mean(fish$INTTIME)
sd(fish$INTTIME)

# summary for funsies
summary(fish$INTTIME)
```

## Q 2 5.56 pg 215

a. mean = 0.21, variance = 0.0147

$$
\alpha = 3 \newline
\beta = 0.07 \newline
\newline
\mu = \alpha\beta = (3)(0.07) = 0.21 \newline
\sigma^{2} = \alpha\beta^{2} = (3)(0.07)^{2} = 0.0147 \newline
$$

b. No, 0.60 falls outside of the 3 sd range when $\alpha=3$ and $\beta=0.07$. The previous gamma distribution no longer models the new maximum flood level distribution for the 4-year period.

$$
\sigma = \sqrt{\sigma^{2}} = \sqrt{0.0147} = 0.1212435565 \newline
\mu \pm 3\sigma = 0.21 \pm 3(0.1212435565) = 0.21 \pm 0.3637306695 = (-0.1537306695, 0.5737306695) \newline
$$

## Q 3 5.60 pg 216

a. meanA = 4, meanB = 4

$$
formula A: \alpha=2, \beta=2 \newline
formula B: \alpha=1, \beta=4 \newline
\mu_A = \alpha\beta=(2)(2)=4 \newline
\mu_B = \alpha\beta=(1)(4)=4 \newline
$$

b. variance A = 8, variance B = 16

$$
\sigma^{2}_{A} = \alpha\beta^{2} = (2)(2)^{2} = 8 \newline
\sigma^{2}_{B} = \alpha\beta^{2} = (1)(4)^{2} = 16 \newline
$$

c.

For A:

$$
P(Y<1)=\int_0^1 \frac{y^{2-1}e^{-y/2}}{2^2\Gamma(2)}dy = \int_0^1 \frac{ye^{-y/2}}{4}dy = \frac{1}{4} \int_0^1 2e^{-y/2}dy + \frac{1}{4} [-2e^{-y/2}\right]_0^1 \newline
$$

$$
= \frac{1}{4} \left[-2e^{-1/2} + 2(0)e^{0}\right] = \frac{1}{4} \left[-2e^{-1/2}(3) + 4\right] = \frac{1}{4} \left[-6e^{-1/2} + 4\right] = \frac{1}{4} \left[-6(0.6065306597) + 4\right] = \frac{1}{4} \left[-3.6391839582 + 4\right] = \frac{1}{4} \left[0.3608160418\right] = 0.09020401045

$$

For B:


