---
title: "lab10"
author: "Tiffany Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
getwd()
```

# Task 2

```{r}
mymaxlik=function(lfun,x,param,...){
  
  # how many param values are there?
  np=length(param)
  # outer -- notice the order, x then param
  # this produces a matrix -- try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
  z=outer(x,param,lfun) # A
  # z is a matrix where each x,param is replaced with the function evaluated at those values
  y=apply(z,2,sum)
  
  # y is a vector made up of the column sums
  # Each y is the log lik for a new parameter value
  plot(param,y,col="Blue",type="l",lwd=2,...)
  # which gives the index for the value of y == max.
  # there could be a max between two values of the parameter, therefore 2 indices
  # the first max will take the larger indice
  i=max(which(y==max(y))) # B
  abline(v=param[i],lwd=2,col="Red")
  
  # plots a nice point where the max lik is
  points(param[i],y[i],pch=19,cex=1.5,col="Black")
  axis(3,param[i],round(param[i],2))
  #check slopes. If it is a max the slope shoud change sign from + to 
  # We should get three + and two -vs
  ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
  return(list(i=i,parami=param[i],yi=y[i],slope=slope))
  
}

```

 z=outer(x,param,lfun) - line A creates matrix z where rows are the assigned items in the "x" parameter, columns are the assigned the items in the "param" parameter, and lfun is the matrix element output (using row and column as parameters)

i=max(which(y==max(y))) - line B finds the index of the max value of y, and given multiple indexes it returns the higher index

Formula for likelihood:

$$
L(p) = \prod_{i=1}^{8}{20\choose y_i} p^{y_i} (1-p)^{20-y_i}
$$

R code for above likelihood is the above mymaxlik()

```{r}

logbin=function(x,param) log(dbinom(x,prob=param,size=10))
bin = mymaxlik(x=c(3, 3, 4, 3, 4, 5, 5, 4), param=seq(0,1,length=1000), lfun=logbin,xlab=expression(p), main="Binomial")

```

# Task 3

```{r}

logpoiss=function(x,param) log(dpois(x,lambda=param)) 
pois = mymaxlik(x=c(4,6,7,6,5), param=seq(0,15,length=1000), lfun=logpoiss, xlab=expression(p), main="Poisson")

```

The algebraic expression for the log likelihood is:

$$
L(x|\lambda) = \prod_{i=1}^{n}f(x|\lambda) = \prod_{i=1}^{n}\frac{e^{-\lambda}\lambda^{x_i}}{x_i!} = \frac{e^{-n\lambda}\lambda^{\sum_{i}x_{i}}}{\prod_{i}x_i!} = \frac{e^{-n\lambda}\lambda^{t}}{\prod_{i}x_{i}!} = e^{-n\lambda}\lambda^{t}
\newline
= l(\lambda)=lnf(x|\lambda)=-n\lambda+tln\lambda
$$

```{r}

myNRML=function(x0,delta=0.001,llik,xrange,parameter="param"){
  f=function(x) (llik(x+delta)-llik(x))/delta
  fdash=function(x) (f(x+delta)-f(x))/delta
  d=1000
  i=0
  x=c()
  y=c()
  x[1]=x0
  y[1]=f(x[1])
  
  while(d > delta & i<100){
    i=i+1
    x[i+1]=x[i]-f(x[i])/fdash(x[i])
    y[i+1]=f(x[i+1])
    d=abs(y[i+1])
  }
  
  layout(matrix(1:2,nr=1,nc=2,byrow=TRUE),width=c(1,2))
  curve(llik(x), xlim=xrange,xlab=parameter,ylab="log Lik",main="Log Lik")
  curve(f(x),xlim=xrange,xaxt="n", xlab=parameter,ylab="derivative",main=  "Newton-Raphson Algorithm \n on the derivative")
  points(x,y,col="Red",pch=19,cex=1.5)
  axis(1,x,round(x,2),las=2)
  abline(h=0,col="Red")
  
  segments(x[1:(i-1)],y[1:(i-1)],x[2:i],rep(0,i-1),col="Blue",lwd=2)
  segments(x[2:i],rep(0,i-1),x[2:i],y[2:i],lwd=0.5,col="Green")
  
  list(x=x,y=y)
}

myNRML(x0=1,delta=0.000001,llik=function(x) log(dpois(12,x)*dpois(10,x)),xrange=c(0,20),parameter="lambda" )

```

the value of $\hat{\lambda}$ as given by the function is 11

# Task 4

```{r}

logbin2=function(theta){log(dbinom(2,prob=theta,size=6)) + log(dbinom(4,prob=theta,size=10))}

mymaxlikg=function(lfun="logbin2",theta) { # default log lik is a combination bin
  nth=length(theta)  # nu. of valuse used in theta
  thmat=matrix(theta,nr=nth,nc=1,byrow=TRUE) # Matrix of theta
  z=apply(thmat,1,lfun) # z holds the log lik values
  zmax=max(which(z==max(z)))  # finding the INDEX of the max lik
  plot(theta,exp(z),type="l") # plot of lik
  abline(v=theta[zmax],col="Blue")   #  verical line through max
  axis(3,theta[zmax],round(theta[zmax],4))  # one tick on the third axis 
  theta[zmax]   # theta corresponding to max lik
}

theta = seq(0, 1, length=1000)
mymaxlikg(lfun="logbin2", theta)

```

p = 0.3753754

# Task 5

$$
p(y_{1},y_{2}|\theta_{1}\theta_{2})=bin(y_{1}|\theta_{1})poiss(y_{2}|\theta_{2}) \newline
log(\theta_{1}\theta_{2})=y_{1}log\theta_{1} + (n-y_{1})log(1-\theta_{1}) + y_{2}log\theta_{2}-\theta_{2}
$$

```{r}

logbinpois=function(theta1,theta2) log(dbinom(4,size=20,prob=theta1))+ log(dpois(4,lambda=theta2))

maxlikg2=function(theta1,theta2,lfun="logbinpois",...){
  n1=length(theta1)
  n2=length(theta2)
  z=outer(theta1,theta2,lfun)
  contour(theta1,theta2,exp(z),...) # exp(z) gives the lik
  maxl=max(exp(z))    # max lik
  coord=which(exp(z)==maxl,arr.ind=TRUE)  # find the co-ords of the max
  th1est=theta1[coord[1]] # mxlik estimate of theta1
  th2est=theta2[coord[2]]
  abline(v=th1est,h=th2est)
  axis(3,th1est,round(th1est,2))
  axis(4,th2est,round(th2est,2),las=1)
  list(th1est=th1est,th2est=th2est)
}

maxlikg2(theta1=seq(0,1,length=1000),theta2=seq(0,10,length=1000))

```

# Task 6

```{r}

mymlnorm=function(x,mu,sig,...){  #x sample vector
  
  nmu=length(mu) # number of values in mu
  nsig=length(sig)
  n=length(x) # sample size
  zz=c()    ## initialize a new vector
  lfun=function(x,m,p) log(dnorm(x,mean=m,sd=p))   # log lik for normal
  
  for(j in 1:nsig){
    z=outer(x,mu,lfun,p=sig[j]) # z a matrix 
    # col 1 of z contains lfun evaluated at each x with first value of mu, 
    # col2 each x with 2nd value of m 
    # all with sig=sig[j]
    y=apply(z,2,sum)
    # y is a vector filled with log lik values, 
    # each with a difft mu and all with the same sig[j]
    zz=cbind(zz,y)
    ## zz is the matrix with each column containing log L values, rows difft mu, cols difft sigmas 
  }
  
  maxl=max(exp(zz))
  coord=which(exp(zz)==maxl,arr.ind=TRUE)
  maxlsig=apply(zz,1,max)
  contour(mu,sig,exp(zz),las=3,xlab=expression(mu),ylab=expression(sigma),axes=TRUE,
  main=expression(paste("L(",mu,",",sigma,")",sep="")),...)
  mlx=round(mean(x),2)  # theoretical
  mly=round(sqrt((n-1)/n)*sd(x),2)
  #axis(1,at=c(0:20,mlx),labels=sort(c(0:20,mlx)))
  #axis(2,at=c(0:20,mly),labels=TRUE)
  abline(v=mean(x),lwd=2,col="Green")
  abline(h=sqrt((n-1)/n)*sd(x),lwd=2,col="Red")
  
  # Now find the estimates from the co-ords
  muest=mu[coord[1]]
  sigest=sig[coord[2]]
  
  abline(v=muest, h=sigest)
  return(list(x=x,coord=coord,maxl=maxl))
  
}

mymlnorm(x=c(10, 12, 13, 15, 12, 11, 10),mu=seq(9,15,length=1000),sig=seq(0.1,4,length=1000))

```

# Task 7

```{r}

mymlbeta=function(x,alpha,beta,...){  #x sample vector
  
  na=length(alpha) # number of values in alpha
  nb=length(beta)
  n=length(x) # sample size
  zz=c()    ## initialize a new vector
  lfun=function(x,a,b) log(dbeta(x,shape1=a,shape2=b))   # log lik for beta
  
  for(j in 1:nb){
    z=outer(x,alpha,lfun,b=beta[j]) # z a matrix 
    # col 1 of z contains lfun evaluated at each x with first value of alpha, 
    # col2 each x with 2nd value of a 
    # all with b=beta[j]
    y=apply(z,2,sum)
    # y is a vector filled with log lik values, 
    # each with a difft alpha and all with the same sig[j]
    zz=cbind(zz,y)
    ## zz is the matrix with each column containing log L values, rows difft alpha, cols difft betas 
  }
  
  maxl=max(exp(zz))    # max lik
  coord=which(exp(zz)==maxl,arr.ind=TRUE)  # find the co-ords of the max
  aest=alpha[coord[1]] # mxlik estimate of alpha
  best=beta[coord[2]]
  contour(alpha,beta,exp(zz),las=3,xlab=expression(alpha),ylab=expression(beta),axes=TRUE,
  main=expression(paste("L(",alpha,",",beta,")",sep="")),...)
  
  abline(v=aest, h=best)
  points(aest,best,pch=19)
  axis(4,best,round(best,2),col="Red")
  axis(3,aest,round(aest,2),col="Red")
  return(list(x=x,coord=coord,maxl=maxl,maxalpha=aest,maxbeta=best))
  
}

set.seed(12)

sam <- rbeta(30, shape1 = 3, shape2 = 4) 

alpha_values <- seq(1, 8, length.out = 100)
beta_values <- seq(2, 10, length.out = 100)
n_resamples <- 12
estimates <- matrix(NA, ncol = 2, nrow = n_resamples)

par(mfrow = c(3, 4), mar = c(4, 4, 2, 2) + 0.1)

# resample and make estimates
for (i in 1:n_resamples) {
  resample <- sample(sam, length(sam), replace = TRUE)
  res <- mymlbeta(resample, alpha_values, beta_values, lwd = 2, labcex = 1)
  estimates[i, ] <- c(res$maxalpha, res$maxbeta)
}

# plot distance between estimates and known pop values
params <- c(3, 4)
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2) + 0.1)
distances <- apply(estimates, 1, function(est) sqrt(sum((est - params)^2)))

plot(distances, type = 'b', pch = 19, xlab = "Resample Index", ylab = "Distance to Known Population Parameters",
     main = "Distances between Estimates and Known Population Parameters")
abline(h = 0, col = "red", lty = 2)

```

# Task 8

```{r}
MATH4753nguy0850::mymlbeta(sample(rbeta(30, shape1 = 3, shape2 = 4) , length(rbeta(30, shape1 = 3, shape2 = 4)), replace = TRUE), seq(1, 8, length.out = 100), seq(2, 10, length.out = 100))
```



