---
title: "ass1"
author: "Tiffany Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Questions done: 15/15

# ---- DATA and FUNCTIONS ----
```{r}
library(Intro2R)
library(ggplot2)
library(dplyr)
library(plotrix)
source("pareto.R")
dird="/Users/tiffanynguyen/Documents/College/Sophomore/Spring_2025/Applied_Statistical_Methods/data/K25936_Downloads/Excel/"
myread=function(csv){
  fl=paste(dird,csv,sep="")
  read.table(fl,header=TRUE,sep=",")
}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
eq <- myread("EARTHQUAKE.csv")
mtbe <- myread("MTBE.csv")
swd <- myread("SWDEFECTS.csv")
volt <- myread("VOLTAGE.csv")
pipe <- myread("ROUGHPIPE.csv")
ant <- myread("GOBIANTS.csv")
gal <- myread("GALAXY2.csv")
```

## Question 1
Final grade will be calculated as follows:

* 15% - 4 assignments (each 3.75%)
* 10% - Laboratories 
* 10% - 2 projects (1st Project 3.33% (10/3), 2nd Project 6.67% (20/3))
* 10% - In-class quizzes
* 5% - Online chapter quizzes
* 20% - Midterm
* 30% - Final

Grade scale:

* A - [90%-100%]
* B - [80%-90%)
* C - [60%-80%)
* D - [50%-60%)
* F - < 50%

How to work out final grades: 

1. Calculate the average grade (in percent) for each category
2. Multiply the average grade by its categorical weight
3. Sum the results from step 2 to get the final grade

## Question 2

### a
```{r}
# d <- read.table("DDT.csv", header=TRUE, sep=",")
r <- rainbow(length(unique(ddt$MILE)))
m=with(ddt, as.numeric(factor(MILE)))
coplot(LENGTH ~ WEIGHT | RIVER * SPECIES,
       data=ddt,
       xlab="Weight",
       ylab="Length",
       col=rep(r, table(ddt$MILE)),
       pch=21
       )

```

### b
```{r}
# y ~ x + I(x^2) : I is "as is" operator and the carrot should be done algebraically
# quadlm <- lm(y ~ x + I(x^2), data)
# summary (quadlm)

m=with(ddt, as.numeric(factor(MILE)))
length(unique(m))
```
The lower left three conditional plots represent CCATFISH in rivers FCM, LCM, and SCM. There seems to be an increasing linear trend/relationship between length and weight.

### c
Line A basically creates a numeric representation of mile for each row in ddt, and maps it accordingly. 

### d
Line B counts the number of unique numeric values from line A 

### e
The top 6 plots are empty because there is no data for LMBASS and SMBUFFALO in rivers FCM, LCM, and SCM.

### f
```{r}
head(ddt)
mean(ddt[ddt$RIVER=="FCM" & ddt$SPECIES=="CCATFISH",]$DDT)

```

## Question 3

a. Quantitative 
b. Quantitative
c. Qualitative
d. Quantitative
e. Qualitative
f. Quantitative
g. Qualitative

## Question 4

#### Simple Random Sample: 
Samples are selected from a population that ensures every subset in the population has an equal chance of being included in the sample. 

#### Stratified Random Sampling: 
Random samples of experimental units are obtained for each strata. Strata are experimental units associated with the population that can be separated into groups.

#### Cluster Sampling: 
Samples are selected from natural groupings/clusters and data is collected from all experimental units within each cluster

#### Systematic Sampling: 
Samples are selected every kth experimental unit from a list of all the experimental units 

## Question 5

```{r}
ind <- sample(1:223, 5, replace=FALSE)
mtbe[ind,]
```

### a
```{r}
# i
mtbeo=na.omit(mtbe)

# ii
sd(mtbeo[mtbeo$Aquifier=="Bedrock",]$Depth)
```

## Question 6

### a
```{r}
# i
eq[sample(1:2929, 30, replace=FALSE),]$MAGNITUDE

# ii
plot(ts(eq$MAGNITUDE))
median(eq$MAGNITUDE)
```

## Question 7

a. Stratified random sampling
b. Fish in the Tennessee River
c. Capture location and species

## Question 8

a. Bar chart
b. Types of Robotic Limbs
c. Legs ONLY
d. 

```{r}
freq=c(15, 8, 63, 20)
RL=c("None", "Both", "Legs Only", "Wheels Only")
table(rep(RL, freq)) / sum(freq)

```

e. 
```{r}
pareto(rep(RL, freq))
```

## Question 9

a. Explorer
```{r}
products = c("Windows" = 32, "Explorer" = 6, "Office" = 12)
pie(products, 
    main="Microsoft Security Issues",
    col=rainbow(length(products))
)
```

b. Remote Code Execution
```{r}
freq = c(6, 8, 22, 3, 11)
categories = c("Denial of Service", "Information Disclosure", "Remote Code Execution", "Spoofing", "Privilege Elevation")
pareto(rep(categories, freq))
```

## Question 10

```{r}
head(swd)
tab=table(swd$defect)
rtab=tab/sum(tab)
round(rtab, 2)
pie3D(rtab, labels=list("OK", "Defective"), main="pie plot of SWD")

```

The module software code is most likely not going to be defective, but we can expect modules to be defective 10% of the time.

## Question 11

### a
```{r}
old<-subset(volt,subset=LOCATION=="OLD")
old$VOLTAGE->vto
lept<-min(vto)-0.05
rept<-max(vto)+0.05
rnge<-rept-lept
inc<-rnge/9
seq(lept, rept,by=inc)->cl
cvto<-cut(vto,breaks=cl)
old.tab=table(cvto)
old.relative.freq <- old.tab / sum(old.tab)

# barplot for old relative frequency
barplot(old.relative.freq,space=0,main="Old Relative Frequency Barplot",las=2, ylab="Relative Frequency")

# histogram for old relative frequency
hist(vto, prob=TRUE, nclass=10, xlab="Voltage", ylab="Relative Frequency", main="Old Relative Frequency Histogram")
```

### b
```{r}
stem((volt %>% filter(LOCATION == "OLD"))$VOLTAGE)
```
The stem-and-leaf plot is more informative about where most of the voltage readings like because its much easier to see where the readings lie.

### c
```{r}
new<-subset(volt,subset=LOCATION=="NEW")
new$VOLTAGE->vtn
lept<-min(vtn)-0.05
rept<-max(vtn)+0.05
rnge<-rept-lept
inc<-rnge/9
seq(lept, rept,by=inc)->cl
cvtn<-cut(vtn,breaks=cl)
new.tab=table(cvtn)
relative.freq <- new.tab / sum(new.tab)

# barplot for new relative frequency
barplot(relative.freq,space=0,main="New Relative Frequency Barplot",las=2, ylab="Relative Frequency")

# histogram for new relative frequency
hist(vtn, prob=TRUE, nclass=10, xlab="Voltage", ylab="Relative Frequency", main="New Relative Frequency Histogram")
```

### d

No, the new process is not as good as the old process. Most of the readings in the old location are above the 9.2/good process line whereas the new location is more dispersed. 

### e
```{r}
mean(vto)
median(vto)
getmode(vto)
mean(vtn)
median(vtn)
getmode(vtn)
```

The preferred measure of central tendency is mean because there are no outliers.

### f
```{r}
(10.5-mean(vto))/sd(vto)
```

### g
```{r}
(10.5-mean(vtn))/sd(vtn)
```

### h

A voltage reading of 10.50 is more likely to occur at the old location because the z-score is lower (less standard deviations from the mean).

### i

```{r}
boxplot((volt %>% filter(LOCATION == "OLD"))$VOLTAGE)
```

There are a few outliers in the old location

### j

```{r}
voltOld = (volt %>% filter(LOCATION == "OLD"))$VOLTAGE
voltOldZ = (voltOld - mean(voltOld)) / sd(voltOld)

voltOld[abs(voltOldZ) > 2]
```
There are 3 outliers in the old location

### k

```{r}
boxplot((volt %>% filter(LOCATION == "NEW"))$VOLTAGE)
```

There seems to be no outliers at the new location

### l

```{r}
voltNew = (volt %>% filter(LOCATION == "NEW"))$VOLTAGE
voltNewZ = (voltNew - mean(voltNew)) / sd(voltNew)

voltNew[abs(voltNewZ) > 2]
```

There are no outliers at the new location

### m
```{r}
boxplot(cbind(voltOld, voltNew))
```

The voltage readings at the old location are higher than the voltage readings at the new location, but the old location has more outliers.


## Question 12

```{r}
r=pipe$ROUGH
mean(r)
sd(r)
```

\begin{eqnarray}

\bar{y} \pm 2s \\
1.881 \pm 2(0.524) \\
1.881 \pm 1.048 &=& (0.833, 2.929) \\

\end{eqnarray}

The interval that will likely contain about 95% of all coated pipe roughness measurements is (0.833, 2.929). 2 standard deviations are taken because approximate 95% of the data will fall within 2 standard deviations of the mean.

## Question 13

### a

```{r}
ant

# average number ant species 
mean(ant$AntSpecies)

# number of ant species discovered at each site is around 5 ish
median(ant$AntSpecies)

# most common number of ant species discovered is 5
getmode(ant$AntSpecies)
```

### b

I would recommend median because there are many outliers.

### c

```{r}
ds <- ant %>% filter(Region == "Dry Steppe")

mean(ds$PlantCov)
median(ds$PlantCov)
getmode(ds$PlantCov)
```

### d

```{r}
gd <- ant %>% filter(Region == "Gobi Desert")

mean(gd$PlantCov)
median(gd$PlantCov)
getmode(gd$PlantCov)
```

### e

Yes, the center of total plant coverage are different in the two regions. Dry Steppe centers around the 40s whereas Gobi Desert centers around the 25-30s.

## Question 14

### a

```{r}
hist(gal$VELOCITY,
     main="Velocity Distribution A1775",
     xlab="Velocity (km/s)")

```


### b

Yes, there are two peaks in the histogram which indicates there is evidence to support the double cluster theory.

### c

I am taking 21000 as the velocity to split the clusters.

```{r}
split = 21000

galA <- gal[gal$VELOCITY < split,]
galB <- gal[gal$VELOCITY >= split,]

# mean for first galaxy
mean(galA)

# standard deviation for first galaxy 
sd(galA)

# mean for second galaxy
mean(galB)

# standard deviation for second galaxy
sd(galB)
```


### d
```{r}
# z scores
(20000-mean(galA))/sd(galA)
(20000-mean(galB))/sd(galB)
```
This galaxy is likely to belong to A1775A because 20,000 km/s is 1.01 sd away from A1775A, compared to 5.06 sd away from A1775B.

## Question 15
```{r}
ggplot(ddt, aes(x=RIVER, y=LENGTH, fill=SPECIES)) +
  geom_boxplot() +
  labs(title="Tiffany Nguyen",
       x="River",
       y="Length")
```
