---
title: 'Assignment 2'
author: "Tiffany Nguyen"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions{17/17}

## Q 1

a.7%
```{r}
# 1 - P(Expert | Match)
1 - 0.9212

```

b. 25.45%
```{r}
# 1 - P(Novice | Match)
1 - 0.7455

```

c. The participant is more likely to be a novice


## Q 2

a. 50% 
```{r}
# In a population of 1,000 athletes, suppose 100 are illegally using testosterone. Of the users, suppose 50 would test positive for testosterone. Of the nonusers, suppose 9 would test positive
# P(Positive | User)
50/100
```

b. 99%
```{r}
# P(negative | nonuser)
(1000 - 100 - 9)/(1000 - 100)
```

c. 84.75%
``` {r}
# If an athlete tests positive for testosterone, use Bayes’ Rule to find the probability that the athlete is really doping.
# P(user | positive)
((50/100) * (100/1000)) / ((50/100) * (100/1000) + (9/900) * (900/1000))
```

## Q 3

3.1 - Informal Proof of the Multiplicative Rule

Step-by-Step (how I prove it in my mind)

1. Choose one element from the first set. You have $n_1$ choices
2. Choose one element from the second set. For each of the previous choice, you have $n_2$ choices, giving a total of  $n_1 n_2$ choices
3. Repeat steps 1 and 2 until you reach the $k^{th}$ set, giving you a total combination choice of $n_1 n_2 n_3 ..... n_k$ choices. This is the number of different samples that can be formed.

## Q 4

3.2 - Informal Proof of the Permutations Rule

There are $N$ ways for the first position, $N - 1$ ways for the second position, $N - 2$ ways for the third position, and so on. The total number of ways to arrange the $n$ objects is $N(N - 1)(N - 2) ... (N - n + 1)$, which is the same as $$\frac{N!}{(N - n)!}$$

## Q 5

3.3 - Informal Proof of the Partitions Rule

Given $N$ elements and $k$ partitions with the first set containing $n_1$ elements, the second set containing $n_2$ elements, and so on, we count the number of ways to arrange $N$ elements in without partitions ($N!$) and divide by the number of ways to arrange the elements within the partitions ($n_1!n_2!...n_k!$). This gives us the formula $$\frac{N!}{n_1!n_2!...n_k!}$$

## Q 6

3.4 - Informal Proof of the Combinations Rule

Selecting $n$ elements from a set of $N$ elements is the same process as partitioning the $N$ elements into two sets: the $n$ elements you want to select and remaining ($N - n$) not selected. Using the partitions rule, we get the formula $$\frac{N!}{n!(N - n)!}$$

## Q 7

a. yes, 1

```{r}
y = 0.09 + 0.3 + 0.37 + 0.2 + 0.04
y
```

b. 24%
```{r}
# P(Y=3) + P(Y=4)
0.2 + 0.04
```

c. 39%
```{r}
# P(Y=0) + P(Y=1)
0.09 + 0.3
```

## Q 8

a. 
Properties of discrete random variables: between 0 and 1, inclusive, and probabilities total up to 1

```{r}
0.17 + 0.1 + 0.11 + 0.11 + 0.1 + 0.1 + 0.07 + 0.05 + 0.03 + 0.02 + 0.02 + 0.02 + 0.02 + 0.02 + 0.01 + 0.01 + 0.01 + 0.01 + 0.01 + 0.005 + 0.005
```

b. 14%

```{r}
# P(Y >= 10)
0.02 + 0.02 + 0.02 + 0.02 + 0.01 + 0.01 + 0.01 + 0.01 + 0.01 + 0.005 + 0.005
```

c. mean: 4.655, variance: 19.8559
```{r}
# mean and variance of Y
mean = 0*0.17 + 1*0.1 + 2*0.11 + 3*0.11 + 4*0.1 + 5*0.1 + 6*0.07 + 7*0.05 + 8*0.03 + 9*0.02 + 10*0.02 + 11*0.02 + 12*0.02 + 13*0.02 + 14*0.01 + 15*0.01 + 16*0.01 + 17*0.01 + 18*0.01 + 19*0.005 + 20*0.005
mean

variance = (0 - mean)^2 * 0.17 + (1 - mean)^2 * 0.1 + (2 - mean)^2 * 0.11 + (3 - mean)^2 * 0.11 + (4 - mean)^2 * 0.1 + (5 - mean)^2 * 0.1 + (6 - mean)^2 * 0.07 + (7 - mean)^2 * 0.05 + (8 - mean)^2 * 0.03 + (9 - mean)^2 * 0.02 + (10 - mean)^2 * 0.02 + (11 - mean)^2 * 0.02 + (12 - mean)^2 * 0.02 + (13 - mean)^2 * 0.02 + (14 - mean)^2 * 0.01 + (15 - mean)^2 * 0.01 + (16 - mean)^2 * 0.01 + (17 - mean)^2 * 0.01 + (18 - mean)^2 * 0.01 + (19 - mean)^2 * 0.005 + (20 - mean)^2 * 0.005
variance
```

d. -4.2570 < Y < 13.5670
```{r}
# give an interval that will contain the value of Y with a probability of at least 0.75

sd = sqrt(variance)
interval <- c(mean - 2*sd, mean+2*sd)
interval
```

## Q 9

a. 0.132%
```{r}
# The National Science Foundation reports that 70% of the U.S. graduate students who earn PhD degrees in engineering are foreign nationals. Consider the number Y of foreign students in a random sample of 25 engineering students who recently earned their PhD.
# P(Y=10)
dbinom(10, 25, 0.7)
```

b. basically 0
```{r}
# P(Y<=5)
pbinom(5, 25, 0.7)
```

c. mean: 17.5, sd: 2.29
```{r}
# mean and standard deviation
mean = 25 * 0.7
mean

sd = sqrt(25 * 0.7 * 0.3)
sd
```

d. In a random sample of 25 engineering students who recently earned their PhD, 17.5 students are expected to be foreign students with a high standard deviation (over 2 sd)

## Q 10

a. basically 0
```{r}
# probability that exactly 5 trains are assigned to each of the 10 tracks
dbinom(5, 50, 0.1)^10
```

b. 3.37%
```{r}
# probability that track 1 is underutilized (<2 trains assigned to the track for the day)
dbinom(0, 50, 0.1) + dbinom(1, 50, 0.1)
```


## Q 11

a. geometric probability distribution
$$
p(y) = (0.60)^{y-1}(0.40) \text{ for } y = 1, 2, 3, ...
$$
b. E(Y) = 2.5
```{r}
# E(Y)
1/0.4
```

c. 40%
```{r}
# P(Y=1)
(0.60)^(1-1)*(0.40)
```

d. 36%
```{r}
# P(Y>2)
# 1 - P(Y=1) - P(Y=2)
1 - ((0.60)^(1-1)*(0.40)) - ((0.60)^(2-1)*(0.40))
```


## Q 12

a. In a sample of 10 facilities, the expected number of facilities that treat hazardous waste on-site is 0.3828
```{r}
8/209*10
```

b. 0.02%
```{r}
dhyper(4,8,201,10)
```

## Q 13

a. 0.03

b. independent events (for casualities), casualities happen at a constant rate over time, events must be discrete, events are non-negative and have no upper-bound

c. 97%
```{r}
# P(Y=0)

dpois(0, 0.03)

```


## Q 14

a. 
$$
\int_{0}^{1} c(2-y)dy = 1 \newline
c(2-\frac{1}{2}) = 1 \newline
c=\frac{2}{3}
$$

b.
$$
 \begin{equation}
 f(x) =
   \left\{\begin{array}{lr}
       \frac{2}{3}(2y-\frac{1}{2}y^2), & 0 \le x \le 1 \\
       0, & elsewhere 
    \end{array}\right.
 \end{equation}
$$
c.
$$
F(0.4) = \frac{2}{3}(2(0.4)-\frac{1}{2}(0.4)^2) = 0.48
$$
d.
$$
P(0.1\le Y\le 0.6) = F(0.6) - F(0.1) \newline
\frac{2}{3}(2(0.6)-\frac{1}{2}(0.6)^2) - \frac{2}{3}(2(0.1)-\frac{1}{2}(0.1)^2) \newline
0.68 - 0.13 = 0.55
$$

## Q 15

a.
$$
\mu=E(Y)\int_{-5}^{5}\frac{3y}{500}(25-y^2)dy \newline
=\int_{-5}^{5}(\frac{3y}{20}-\frac{3y^3}{500})dy =(\frac{3y^2}{40}-\frac{3y^4}{2000})|_{-5}^{5} \newline
= (\frac{3(5)^2}{40}-\frac{3(5)^4}{2000})-(\frac{3(-5)^2}{40}-\frac{3(-5)^4}{2000}) \newline\
=(1.875-0.9375)-(1.875-0.9375) = 0
$$
$$
E(Y^2)=\int_{-5}^{5}y^2\frac{3y^2}{500}(25-y^2)dy \newline
=\int_{-5}^{5}(\frac{3y^2}{20}-\frac{3y^4}{500})dy \newline
=\frac{y^3}{20}-\frac{3y^5}{2500}|_{-5}^{5} \newline
= (\frac{(5)^3}{20}-\frac{3(5)^5}{2500})-(\frac{(-5)^3}{20}-\frac{3(-5)^5}{2500}) \newline
=(6.25-3.75)-(-6.25+3.75)=5 \newline
\sigma^2=E(Y^2)-\mu^2=5-0^2=5
$$
b. 
$$
\text {Let }X=\frac{1}{60}Y. \text { Then } \mu_x=E(X)=E(\frac{1}{60}Y)=\frac{1}{60}E(Y)=\frac{1}{60}(0)=0 \newline
\sigma_x^2=\sigma_{1/60Y}^2=(\frac{1}{60})^2\sigma_Y^2=\frac{1}{3600}(5)=0.0014
$$

c.
$$
\text {Let }Z=60Y. \text { Then }\mu_z=E(Z)=E(60Y)=60E(Y)=60(0)=0 \newline
\sigma_z^2=\sigma_{60Y}^2=(60)^2\sigma_Y^2=3600(5)=18,000
$$

## Q 16

a. 94%
```{r}
# mean of 50 mg per liter, sd of 3.2 mg per liter
# P(Y>45)
1 - pnorm(45, 50, 3.2)
```

b. 94%
```{r}
# P(Y<55)
pnorm(55, 50, 3.2)
```

c. 11%
```{r}
# P(51<Y<52)
pnorm(52, 50, 3.2) - pnorm(51, 50, 3.2)
```

## Q 17

a. 41%
```{r}
# mean of 605 points, sd of 185 points
# P(500<y<700)
pnorm(700, 605, 185) - pnorm(500, 605, 185)
```

b. 15%
```{r}
# P(400<Y<500)
pnorm(500, 605, 185) - pnorm(400, 605, 185)
```

c. 91%
```{r}
# P(Y<850)
pnorm(850, 605, 185)
```

d. 2%
```{r}
# P(Y>1000)
1 - pnorm(1000, 605, 185)
```

e. 842 points
```{r}
# what rating will only 10% of the crash-tested cars exceed?
qnorm(0.9, 605, 185)
```




